---
title: "Scrape Strava Data"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
#..........................load packages.........................
library(rStrava) 
library(tidyverse)
library(lubridate)
library(leaflet)
library(googleway)
library(htmlwidgets)
library(htmltools)
library(aws.s3)
```

## Setup

```{r}
#.......Strava app name, client ID, secret, and athlete id.......
app_name <- Sys.getenv("STRAVA_APP_NAME")
app_client_id <- Sys.getenv("STRAVA_KEY")
app_secret <- Sys.getenv("STRAVA_SECRET")

#......................create strava token.......................
my_token <- httr::config(token = strava_oauth(app_name, app_client_id, app_secret,
                                              app_scope = "activity:read_all",
                                              cache = TRUE))
```

## Scrape data

```{r}
#....................scrape strava activities....................
my_activities <- rStrava::get_activity_list(stoken = my_token)

#......................number of activities......................
length(my_activities) 
```

## Wrangle data

```{r}
#...............compile and clean up activity data...............
my_acts <- rStrava::compile_activities(my_activities) |> 
  
  # convert units
  mutate(total_miles = round(distance * 0.62137, digits = 2), # convert km to miles
         elevation_gain_ft = round(total_elevation_gain * 3.28084, digits = 2), # convert meters to ft
         elev_high_ft = round(elev_high * 3.28084, digits = 2), # convert meters to ft
         elev_low_ft = round(elev_low * 3.28084, digits = 2), # convert meters to ft
         avg_speed_mi_hr = round(average_speed/1.609, digits = 2), # convert km/hr to mi/hr 
         max_speed_mi_hr = round(max_speed/1.609, digits = 2), # convert km/hr to mi/hr 
         moving_time_hrs_mins = seconds_to_period(moving_time), # convert sec to hr/min/sec
         elapsed_time_hrs_mins = seconds_to_period(elapsed_time)) |> # convert sec to hr/min/sec
  
  # convert start_date_local to Date object
  mutate(start_date_time_local = as_datetime(x = start_date_local),
         start_date_local = as_date(x = start_date_local),
         start_date = as_datetime(x = start_date),
         week_number = isoweek(start_date_local)) |> 
  
  # separate date/time components into different columns
  mutate(year = year(start_date_local), 
         month = month(start_date_local), 
         day = day(start_date_local), 
         hour = hour(start_date_local),
         minute = minute(start_date_local),
         second = second(start_date_local)) |> 
  
  # convert character vars to factors 
  mutate_at(c("visibility", "sport_type", "has_heartrate", 
              "gear_id", "athlete.id", 
              "upload_id_str", "private"), 
            factor) |> 
  
  # add column of alternate sport_type names
  mutate(sport_type_alt = case_when(
    sport_type == "Hike" ~ "Hike",
    sport_type == "Ride" ~ "Bike",
    sport_type == "GravelRide" ~ "Bike",
    sport_type == "Walk" ~ "Walk"
  )) |> 
  
  # add gear name based on gear_id
  mutate(gear_name = case_when(
    gear_id == "g8567330" ~ "danner1",
    gear_id == "g8567301" ~ "danner2",
    gear_id == "g9707381" ~ "danner3",
    gear_id == "g12612884" ~ "danner4",
    gear_id == "g10199228" ~ "lowa",
    gear_id == "g14316125" ~ "hoka1",
    gear_id == "g18225795" ~ "hoka2",
    gear_id == "g18258672" ~ "tevas",
    gear_id == "g10257186" ~ "nano",
    gear_id == "b9318843" ~ "grail",
    gear_id == "b9263591" ~ "tandem",
    gear_id == "b9263606" ~ "yeti",
    gear_id == "b9263624" ~ "trek",
    gear_id == "b9263626" ~ "cannondale"
  )) |> 
  
  # filter out only_me visibility
  filter(visibility != "only_me") |> 

  # remove columns that were mutated above  
  dplyr::select(-c(distance, total_elevation_gain)) |> 
  
  # select, rename & reorder columns
  dplyr::select(
    # activity & athlete info
    id, name, visibility, sport_type, sport_type_alt, start_date_time_local, start_date_local,
    year, month, day, hour, minute, second, week_number, timezone, 
    athlete_count, achievement_count, pr_count,
    # activity stats
    total_miles, elevation_gain_ft, 
    avg_speed_mi_hr, max_speed_mi_hr, 
    moving_time_hrs_mins, moving_time_sec = moving_time, 
    elapsed_time_hrs_mins, elapsed_time_sec = elapsed_time, 
    elev_high_ft, elev_low_ft,
    # map info
    map.id, map.resource_state, map.summary_polyline, 
    lat = start_latlng1, lng = start_latlng2, end_latlng1, end_latlng2,
    # other vars
    has_heartrate, average_heartrate, gear_id, gear_name, max_heartrate,
    # interactions
    kudos_count, comment_count,
    # other metadata
    athlete.id, upload_id, upload_id_str, external_id, id, private, start_date, utc_offset
    ) |> 

  # filter out any obs that we don't want to include
  filter(id != "5502812467", # remove hike that didn't recorded correctly
         !sport_type %in% c("Workout", "Kayaking")) # remove recorded workouts (don't have location data associated with them)

#...................fix incorrect week numbers...................
y2020 <- my_acts |> filter(year == "2020")

y2021 <- my_acts |> filter(year == "2021") # jan 1 needs to be updated to week_number 1
y2021[109, 14] = 1

y2022 <- my_acts |> filter(year == "2022") # jan 1 needs to be updated to week_number 1
y2022[104, 14] = 1

y2023 <- my_acts |> filter(year == "2023") # jan 1 needs to be updated to week_number 1
y2023[101, 14] = 1

y2024 <- my_acts |> filter(year == "2024") # no jan 1 hike, so no updates needed

#..............add 0s during weeks with no activity..............
#tbd

#..........................recombine dfs.........................
wrangled_activities <- rbind(y2024, y2023, y2022, y2021, y2020)

#............................save data...........................
saveRDS(wrangled_activities, here::here("strava_dashboard/data/wrangled_activities.rds"))

# s3saveRDS(
#   x = wrangled_activities,
#   bucket = "sams-strava-dashboard-data"
# )
```

The above processed data does not contain the following variables (they were removed because they felt unnecessary to keep, at least for now, and/or it was unclear what they represented):

-   `commute`
-   `display_hide_heartrate_option`
-   `flagged`
-   `from_accepted_tag`
-   `heartrate_opt_out`
-   `device_watts`
-   `average_watts`
-   `athlete.resource_state`
-   `trainer`
-   `photo_count`
-   `total_photo_count`
-   `has_kudoed`
-   `manual`
-   `resource_state`
-   `workout_type`
-   `kilojoules`
